Improved Quasi-Static Methods for Time-Dependent Neutron Diffusion and Implementation in Rattlesnake

### 2017-03-05 doc/Defense/Defense_Prince.tex
Defense: Initial draft (complete)

\usepackage{tikz}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{multicol}

@@

\begin{block}{Purpose Statement}
Investigate and develop the improved quasi-static method (IQS) for minimizing computation expense for transient reactor simulations, while maintaining accuracy.
\end{block}

@@

@@

@@

@@

@@

@@

>>>>

\usepackage{tikz}
\usepackage{colortbl}
%\usepackage{booktabs}
\usepackage{multicol}
\usepackage{caption}

@@

\begin{block}{Purpose Statement}
Investigate and develop the improved quasi-static method (IQS) for minimizing computation expense in transient reactor simulations, while maintaining accuracy.
\end{block}

@@

%-------------------------------------------------------------------
\begin{frame}{One Dimensional Example}

\begin{block}{Geometry}
\begin{figure}
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}}
\end{center}
\end{figure}
\begin{itemize}
\item Region 2: Ramp $\Sigma_a$ decrease (rod removal)
\item Region 3: Ramp $\Sigma_a$ decrease then increase (rod removal then insertion)
\item Region 4: Ramp $\Sigma_a$ increase (rod insertion)
\end{itemize}
\end{block}

\begin{block}{Region Perturbations}
\centering
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Baseline Results}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.45\linewidth]{figures/1D_flux.png}
\includegraphics[width=0.55\linewidth]{figures/1D_power_base.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{IQS Results}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.45\linewidth]{figures/1D_shape2.png}
\includegraphics[width=0.55\linewidth]{figures/1D_iqs_power.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{Convergence criteria convergence behavior}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_renorm.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{IQS converged error with theta integration of precursors}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_error.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{IQS converged error with analytical integration of precursors}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_error_an.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{First and second order time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_IE.png}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF2.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{Third and fourth order time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF3.png}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF4.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{SDIRK33 time step error convergence}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/1D_conv_SDIRK33.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{Rattlesnake time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_Rat.png}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_lin.png}
\end{figure}
\end{columns}

@@

%-------------------------------------------------------------------
\begin{frame}{TWIGL Benchmark}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/twigl_geom.png}
\includegraphics[width=0.5\linewidth]{figures/ndiff_ramp_flux.jpg}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{TWIGL Power Profile}

\begin{columns}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot.png}
\caption{TWIGL Power Profile}
\end{figure}

\column{.55\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot2.png}
\caption{TWIGL Peak Power Profile}
\end{figure}

\end{columns}

%-------------------------------------------------------------------
\begin{frame}{TWIGL time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/TWIGL_convergence.jpg}
\includegraphics[width=0.5\linewidth]{figures/TWIGL_ramp_lin.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{TWIGL with Time Adaptation}

\begin{columns}

\column{.7\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot_dt2.png}
\caption{TWIGL Power Profile}
\end{figure}

\column{.4\textwidth}
\begin{table}
\begin{center}
\colorbox{white}{\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|}
\hline
\multicolumn{5}{|c|}{Implicit Discretization}\\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9    &	29 \\
2 &	0.01    &	3.5555e-05 &	11   &	35 \\
3 &	0.005   &	4.0364e-05 &	11   &	31  \\
4 &	0.001   &	0.00294822 &	33   &	122  \\
5 &	0.0005  &	0.00099778 &	39   &	131  \\
\hline
\rowcolor{yellow} 6 &	0.0001  &	0.00019510 &	78   &	236  \\
\hline
7 &	5.0e-05 &	0.00018372 &	112  &	342  \\
8 &	1.0e-05 &	8.0564e-05 &	263  &	794 \\
\hline

\multicolumn{5}{|c|}{IQS} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05 	&	0.03380433 &	4   &	20   \\
2 &	0.01 	&	0.00166991 &	5   &	40  \\
3 &	0.005  	&	0.00886584 &	5   &	40  \\
4 &	0.001 	&	0.02976305 &	5   &	36  \\
5 &	0.0005 	&	0.00143781 &	6   &	55  \\
\hline
\rowcolor{yellow} 6 &	0.0001 	&	0.00016175 &	8   &	65  \\
\hline
7 &	5.0e-05 &	6.0328e-05 &	12  &	163  \\
8 &	1.0e-05 &	7.7103e-05 &	379 &	5729 \\
\hline

\multicolumn{5}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.03380433 &	4  &	9   \\
2 &	0.01    &	0.00263068 &	5  &	12  \\
3 &	0.005   &	0.00160486 &	6  &	21  \\
\hline
\rowcolor{yellow} 4 &	0.001   &	1.7527e-05 &	10 &	35  \\
\hline
5 &	0.0005  &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05  &	9.8321e-07 &	48 &	210 \\
\hline
\end{tabular}}}
\end{center}
\end{table}

\end{columns}

@@

%-------------------------------------------------------------------
\begin{frame}{LRA Benchmark}
\begin{columns}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lra.png}
\end{figure}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lra_profile2D.png}
\end{figure}

\end{columns}
\end{frame}
%-------------------------------------------------------------------

%-------------------------------------------------------------------
\begin{frame}{LRA Power and Temperature Profile}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/lra_profile.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{LRA Time Step Error Convergence}
\begin{columns}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lra_bad.png}
\caption{Only one temperature update per macro step}
\end{figure}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/lra_mp_convergence.png}
\caption{Five temperature updates per macro step}
\end{figure}

\end{columns}
\end{frame}
%-------------------------------------------------------------------

%-------------------------------------------------------------------
\begin{frame}{Analysis on Temperature Updates}

\begin{columns}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{figures/lra_mp.png}
\end{figure}

\column{.5\textwidth}
\begin{table}
\captionsetup{font=scriptsize}
\begin{center}
\colorbox{white}{\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
\multicolumn{5}{|c|}{Implicit Discretization}\\
\hline
Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
\hline
1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
4 	& 5.0e-4 	& 1.892e-4 	& 21.91	& 2.08e5	\\
5 	& 2.5e-4	& 4.590e-5 	& 25.23	& 3.16e5	\\
\hline
\end{tabular}}}
\end{center}
\end{table}
\vspace{-7mm}

\begin{table}
\captionsetup{font=small}
\begin{center}
\colorbox{white}{\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
\multicolumn{5}{|c|}{IQS}\\
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
3 	& 4 	& 5.796e-4 	& 7.87	&  92.3\%	\\
4 	& 8 	& 4.772e-4 	& 12.61	& 207.9\% 	\\
5 	& 16	& 4.516e-4 	& 22.14	& 440.7\%	\\
\hline
\end{tabular}}}
\end{center}
\end{table}
\vspace{-7mm}

\begin{table}
\captionsetup{font=scriptsize}
\begin{center}
\colorbox{white}{\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
\multicolumn{5}{|c|}{IQS P-C}\\
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 3.488e-3 	& 2.91 	& -28.9\%	\\
2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
5 	& 16	& 7.905e-4 	& 8.19	&  100\%	\\
\hline
\end{tabular}}}
%}
\\
\tiny $^*$ runtime difference from $\Delta t = 0.004$ implicit dis.
\end{center}
\end{table}

\end{columns}

%-------------------------------------------------------------------
\begin{frame}{LRA Dynamical Time Scale Analysis}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/time_constant_lra.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{LRA with Time Adaptation}

\begin{columns}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/LRA_DT2.png}
\caption{LRA Power Profile}
\end{figure}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/LRA_DT2_peak.png}
\caption{LRA Peak Power Profile}
\end{figure}

\end{columns}

\begin{table}
\begin{center}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Event & Power (W/cm$^3$) & Error & Steps & Power (W/cm$^3$) & Error & Steps \\
\hline
Max Power & 5567.3 & 0.019454 & 423 & 5568.3 & 0.019274 & 47 \\
End (3 s) & 109.66 & 2.3650e-4 & 603 & 109.65 & 3.0622e-4 & 97 \\
\hline
\end{tabular}}
\end{center}
\end{table}

@@

%-------------------------------------------------------------------
\begin{frame}{TREAT: Transient-15 Example}
\begin{columns}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_config.png}
\end{figure}

\column{0.5\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_mesh_homo.png}
\end{figure}

\end{columns}
\end{frame}
%-------------------------------------------------------------------

%-------------------------------------------------------------------
\begin{frame}{Transient-15 Power Profile}
\begin{columns}

\column{0.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_core2.png}
\end{figure}

\column{0.5\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_profile.png}
\end{figure}

\end{columns}
\end{frame}
%-------------------------------------------------------------------

%-------------------------------------------------------------------
\begin{frame}{Transient-15 Results}

\begin{block}{Accuracy and Runtime}
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|ccccccc|}
\hline
 & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
\hline
Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
Implicit Dis. 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
\hline
\end{tabular}
}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from implicit discretization 
\end{block}

\begin{block}{Dynamical Time Scale}
\begin{figure}
\includegraphics[height=1.5in]{figures/time_constant_tran15.png}
\end{figure}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Transient-15 with Time Adaptation}

\begin{columns}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_DT2.png}
\caption{TREAT Power Profile}
\end{figure}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_DT2_peak.png}
\caption{TREAT Peak Power Profile}
\end{figure}

\end{columns}

@@

%-------------------------------------------------------------------
\begin{frame}{Conclusions}

\begin{block}{IQS Iteration}
IQS was testing with fixed point itertion used the following criteria:
\begin{enumerate}
\item $L^\infty$ norm of the change in shape between iterations
\item $L^2$ norm of the change in shape between iterations
\item Difference in reactivity between iterations
\item Difference in amplitude between iterations
\item IQS uniqueness consistency criteria
\end{enumerate}
\begin{itemize}
\item 1-4 had equivalent behavior
\item 5 required analytical integration of precursors for reasonable error convergence
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Conclusions (cont.)}

\begin{block}{IQS Time Step Error Convergence}
\begin{itemize}
\item Demonstrated proper convergence through order 4
\item Required higher order interpolation of PRKE parameters for higher order convergence
\item Effectiveness of error convergence demonstrated through time adaptation
\item Marginal improvement over implicit discretization for "difficult" 1D problem
\item Significant improvement for TWIGL and LRA
\end{itemize}
\end{block}

\begin{block}{IQS with Temperature Feedback}
\begin{itemize}
\item IQS was successfully implemented into the Rattlesnake/MOOSE framework
\item Incorporation of temperature into the quasi-static process significantly improved computational efficiency
\item Dynamical time scale analysis demonstrated proof-of-concept for variant time scales
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Recommendations for Future Research}

\begin{block}{Multiphysics}
Test IQS with other reactor physics, including:
\begin{itemize}
\item Thermal Hydraulics
\item Fuel Mechanics
\item Depletion
\end{itemize}
\end{block}

\begin{block}{Transport}
Apply IQS to other transport models (other Rattlesnake action systems):
\begin{itemize}
\item DFEM-Diffusion
\item SAAF-Sn
\item SAAF-Pn
\item LS-Sn
\end{itemize}
\end{block}

\begin{block}{PJFNK}
\begin{itemize}
\item Investigate PJFNK iteration of IQS
\item Optimize IQS preconditioning
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Questions?}

\centering
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/question_mark.jpg}

@@

\begin{figure}
\includegraphics[height=0.5in]{figures/thank_you.png}
\end{figure}

\begin{figure}
\includegraphics[height=2.5in]{figures/nuclear_collage.png}
\hspace{5mm}
\includegraphics[height=2.5in]{figures/tamu_atom.jpg}
\end{figure}
